---
title: "AlgDesign_Tutorial"
author: "Ijeamaka Anyene"
date: "12/18/2019"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library(AlgDesign)
```

### Creating Experimental Design  
When creating an incomplete block design for Max Diff the rules of thumb are as follows:  
    +  Ideal Number of Questions = 3 * Alternatives / Alternatives Per Quesion  
    + Alternatives per question = 3 to 6 (4 or 5 ideal)  
    
```{r}
alternatives = 8 # what are the different products/attributes/treatments that need to be shown?
block_size = 4 # alternatives per question
blocks = (3 * alternatives) / block_size # number of questions
```

Using the `AlgDesign` package - we will create an incomplete block design. AlgDesign - will create a block design based on the parameters you pass - whether it is balanced or not. There is an element of randomness with AlgDesign, so you need to set the seed for reproducibility. 

```{r}
set.seed(2)
BIB = optBlock(~., withinData = factor(1:alternatives), blocksizes = rep(block_size, blocks))
```

BIB returns a few things:  
    + diagonality which is an element of assessing how good the design is (but questionable how useful it is)  
    + Blocks: which is each block as a different item in a list  
    + Design: The order in which the alternatives will appear in your design  
    
Unfortunately, BIB doesn't return the design in the order we would personally prefer to review the design. Here is some code to easily modify it!

```{r}
block_design_matrix = matrix(data = NA, nrow = block_size, ncol = blocks)
n = length(BIB$Blocks)

for(i in 1:n){
  temp = BIB$Blocks[[i]][,1]
  block_design_matrix[,i] = temp
}

block_design_matrix = t(block_design_matrix)
block_design_matrix # Experimental Design 
```


### Creating Count Matrix Based off of Experimental Design
One of the most important assessments of a design is to understand how many times one alternative appears with another alternative. For example, in the above experimental design alternative 1 shows up with alternative 2 - 2 different times. It can be pretty manual to calculate this - so below is some code to create a "count matrix". 


```{r}
# Code: Rep range is the number of rows in the experimental design, each is the number of columns
crossprod(table(rep(1:6, each = 4), unlist(BIB$design)))
```

We can see in the above that this isn't a particularly balanced design. 
    + Alternative 3 never appears with Alternative 7. (Bad)
    + All the alternatives appear at least three times - via the diagnol (Good)
    + Some alternatives appear with the others only 1 or 2 times (Bad)
    
Say we wanted to work with this design anyway! For any analysis using mlogit - you will need to convert the `block_design_matrix` into a matrix with dummy variables. Below is some code to do so!  
  
### Turning Experimental Design into Dummy Variables for Analysis 
```{r}
experimental_design = as.vector(t(block_design_matrix))
sets = rep(1:8, each = 3)
```


```{r}
profiles = model.matrix(~ experimental_design + sets)[, -1] # Removes the intercept
profiles = as.character(profiles) # Convert to characters for the following code
profiles_coded = fastDummies::dummy_cols(profiles)
profiles_coded = apply(profiles_coded, 2, as.numeric)
```

```{r}
head(profiles_coded)
```

This is almost what we want! However, as a reminder max diff designs require 1 for what is "most" and -1 for what is "least". The design above does not have any negatives ones. This will be resolved next. 


```{r}
rows = seq(1, 48, by = 4) # Because each new block starts every 4 rows
profiles_coded_final = matrix(ncol = 9, nrow = 1) 

for(i in rows){
  temp = profiles_coded[i:(i+3), ]
  temp = rbind(temp, -temp) # combining the same set with positive version and negative version 
  
  profiles_coded_final = rbind(profiles_coded_final, temp)
  
  }

profiles_coded_final = profiles_coded_final[-1, ]
```


```{r}
head(profiles_coded_final)
```
